---
title: "R Notebook"
output: html_notebook
---
```{r}
library(plotly)
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggmap)
library(maps)
library(rmarkdown)
library(knitr)
library(viridis)
library(gridExtra)
library(stringr)
setwd("/Users/zackchou/Documents/BAData")
getwd()
```

## Data Manipulation
```{r}
load(file = "/Users/zackchou/Desktop/calldata.rda")
load(file = "/Users/zackchou/Desktop/calldata_2.rda")
load(file = "/Users/zackchou/Desktop/zipcode_map.rda")

call1 <- call1 %>%
  mutate(Call.Type = ifelse(Call.Resolution  %in% c("Gave Caller Information"), "Given Info", 
                            ifelse(Call.Resolution  %in% c("Caller Hung Up", "Got Voicemail (City)", "Info Not Available (Non-City)", "Line Busy (City)", "N/A", "Static/Ghost Call"), "Invalid Info",
                                   ifelse(Call.Resolution  %in% c("Referred To 411", "Referred To County", "Referred To Other Governmental", "Referred To State"), "Referred Info",
                                          ifelse(Call.Resolution  %in% c("Transfer (City)", "Transferred To 411", "Warm Transfer (City)"), "Transferred Info",
                                                 ifelse(Call.Resolution  %in% c("Service Request Processed"),"Processed Info", "None"))))))
save(call1, file = "calldata.rda")

str(newdata)
newdata$CreatedDate <- mdy_hms(newdata$CreatedDate)
newdata$UpdatedDate <- mdy_hms(newdata$UpdatedDate)
newdata <- newdata %>% 
        mutate(wd = wday(CreatedDate, label = T),
               mth = month(CreatedDate, label = T),
               hr = hour(CreatedDate),
               yr = year(CreatedDate))
newdata$ZipCode <- as.numeric(as.character(newdata$ZipCode))
newdata <- newdata %>% 
  select(-c(SRNumber, ActionTaken, Owner, Status, Anonymous, AssignTo, ServiceDate, ClosedDate, AddressVerified, ApproximateAddress, Address, HouseNumber, StreetName, Suffix, TBMPage, TBMColumn, TBMRow, APC, CD, CDMember, NC, NCName, PolicePrecinct))

save(newdata, file = "calldata_2.rda")

zipcode_map$id <- as.numeric(zipcode_map$id)
save(zipcode_map, file = "zip_boundary.rda")

lazip <- zipcode_map %>% 
  filter(city == "Los Angeles")
save(lazip, file = "la_zip.rda")
```

## Yearly Time Distribution 2011-2015
```{r}
k <- call1 %>% 
        filter(!is.na(yr)) %>% 
        group_by(yr, wd, hr) %>% 
        summarise(count = n()) %>% 
        ggplot(aes(x = wd, y = factor(hr), fill = count))+
        geom_tile()+
        scale_fill_viridis()+
        labs(title = "Yearly Call Time Distribution 2011-2015", y = "Hour", x = "Weekday", fill = "Frequency")+
        theme_classic()+
        facet_wrap(~yr)+
        theme(panel.background = element_rect(color = "black"), 
              axis.text.x = element_text(angle = 45, vjust = 0.5),
              axis.text.y = element_text(size = 6))

ggsave("k.png",path = "/Users/zackchou/Desktop/project _results")

```

## Monthly Time Distribution 2011-2015
```{r}
l <- call1 %>% 
        filter(!is.na(yr)) %>% 
        group_by(mth, wd, hr) %>% 
        summarise(count = n()) %>% 
        ggplot(aes(x = wd, y = factor(hr), fill = count))+
        geom_tile()+
        scale_fill_viridis()+
        labs(title = "Montly Call Time Distribution 2011-2015", y = "Hour", x = "Weekday", fill = "Frequency")+
        theme_classic()+
        facet_wrap(~mth)+
        theme(panel.background = element_rect(color = "black"), 
              axis.text.x = element_text(angle = 45, vjust = 0.5),
              axis.text.y = element_text(size = 4))
ggsave("l.png",path = "/Users/zackchou/Desktop/project _results")

```

## Yearly Spatial Distribution 2011-2015
```{r}

la <- qmap("Mid City LA", maptype = "road", zoom = 11, color = "bw")

data1 <- call1 %>% 
  filter(!is.na(Zip.Code),
         !Zip.Code %in% c(99999),
         Service.Name != "") %>% 
  group_by(yr, mth, Zip.Code, Service.Name) %>% 
  summarise(count = n()) %>% 
  arrange(-count)

joindata1 <- left_join(lazip, data1, by = c("id" = "Zip.Code"))

m <- la+
  geom_polygon(data = joindata1, 
               aes(x = long, y = lat, group = group, fill = count, alpha = count),
               color = "darkgrey")+
  theme_void()+
  scale_fill_viridis()+
  labs(title = "Yearly Call Frequency by Zip Area 2011-2015", fill =  "Frequency", alpha = "Transparency")+
  facet_wrap(~yr)

ggsave("m.png",path = "/Users/zackchou/Desktop/project _results")

```

## Monthly Spatial Distribution 2011-2015
```{r}
n <- la+
  geom_polygon(data = joindata1, 
               aes(x = long, y = lat, group = group, fill = count, alpha = count),
               color = "darkgrey")+
  theme_void()+
  scale_fill_viridis()+
  labs(title = "Monthly Call Frequency by Zip Area 2011-2015", fill =  "Frequency", alpha = "Transparency")+
  facet_wrap(~mth)

ggsave("n.png",path = "/Users/zackchou/Desktop/project _results")

```

## Yearly Time Distribution 2016
```{r}
o <- newdata %>% 
        filter(!is.na(yr),
               yr == 2016) %>% 
        group_by(yr, wd, hr) %>% 
        summarise(count = n()) %>% 
        ggplot(aes(x = wd, y = factor(hr), fill = count))+
        geom_tile()+
        scale_fill_viridis()+
        labs(title = "Yearly Call Time Distribution 2016", y = "Hour", x = "Weekday", fill = "Frequency")+
        theme_classic()+
        facet_wrap(~yr)+
        theme(panel.background = element_rect(color = "black"), 
              axis.text.x = element_text(angle = 45, vjust = 0.5))
ggsave("o.png",path = "/Users/zackchou/Desktop/project _results")

```

## Monthly Time Distribution 2016
```{r}
p <- newdata %>% 
        filter(!is.na(yr),
               yr == 2016) %>% 
        group_by(mth, wd, hr) %>% 
        summarise(count = n()) %>% 
        ggplot(aes(x = wd, y = factor(hr), fill = count))+
        geom_tile()+
        scale_fill_viridis()+
        labs(title = "Monthly Call Time Distribution 2016", y = "Hour", x = "Weekday", fill = "Frequency")+
        theme_classic()+
        facet_wrap(~mth)+
        theme(panel.background = element_rect(color = "black"), 
              axis.text.x = element_text(angle = 45, vjust = 0.5))
ggsave("p.png",path = "/Users/zackchou/Desktop/project _results")

```

## Yearly Spatial Distribution 2016
```{r}
data2 <- newdata %>% 
  group_by(yr, mth, ZipCode, RequestType) %>% 
  summarise(count = n()) %>% 
  arrange(-count)

joindata2 <- left_join(lazip, data2, by = c("id" = "ZipCode"))

la <- qmap("2000 Wellington Rd Los Angeles, CA 90016", maptype = "road", zoom = 11, color = "bw")

q <- la+
  geom_polygon(data = filter(joindata2, yr == 2016), 
               aes(x = long, y = lat, group = group, fill = count, alpha = count),
               color = "darkgrey")+
  theme_void()+
  scale_fill_viridis()+
  labs(title = "Call Frequency by Zip Area 2016", fill =  "Frequency", alpha = "Transparency")

ggsave("q.png",path = "/Users/zackchou/Desktop/project _results")

```

## Monthly Spatial Distribution 2016
```{r}
r <- la+
  geom_polygon(data = filter(joindata2, yr == 2016), 
               aes(x = long, y = lat, group = group, fill = count, alpha = count),
               color = "darkgrey")+
  theme_void()+
  scale_fill_viridis()+
  labs(title = "Monthly Call Frequency by Zip Area 2016", fill =  "Frequency", alpha = "Transparency")+
  facet_wrap(~mth)
ggsave("r.png",path = "/Users/zackchou/Desktop/project _results")

```

## Efficiency
```{r}
newdata <- newdata %>% 
  mutate(proc_time2 = UpdatedDate - CreatedDate,
         detector = ifelse(proc_time2 == 0, "T", "F")) 

newdata$proc_time2 <- round(newdata$proc_time2/(60*60*24), digits = 0)

newdata %>% 
  filter(detector != "T") %>% 
  group_by(RequestType) %>% 
  summarise(Average = mean(proc_time2)) %>% 
  arrange(-Average)
```

## Average efficiency by request type
```{r}
a <- newdata %>% 
  group_by(RequestType) %>% 
  summarise(Average = mean(proc_time)) %>% 
  ggplot(aes(x = reorder(RequestType,Average), y = Average, fill = Average))+
  geom_bar(stat = "identity", show.legend = F)+
  coord_flip()+
  labs(title = "Procedure Time", x = NULL, y = "Avg. Days")+
  scale_fill_viridis()+
  theme_classic()
```

## Total number of request of each type
```{r}
levels(newdata$RequestType)
newdata$RequestType = factor(newdata$RequestType, 
                                      levels = c("Report Water Waste",
                                      "Dead Animal Removal",
                                      "Graffiti Removal",
                                      "Metal/Household Appliances",
                                      "Electronic Waste",
                                      "Bulky Items",
                                      "Illegal Dumping Pickup",
                                      "Single Streetlight Issue",
                                      "Other",
                                      "Feedback",
                                      "Multiple Streetlight Issue",
                                      "Homeless Encampment"))


b <- newdata %>% 
  group_by(RequestType) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = RequestType, y = count/1000, fill = RequestType))+
  geom_bar(stat = "identity", show.legend = F)+
  coord_flip()+
  labs(title = "Calling Number Count", x = NULL, y = "Count (in Thousands)")+
  theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis(discrete = T)
```

## Comparison Chart
```{r}
s <- grid.arrange(a, b, nrow = 1)
ggsave("s.png",path = "/Users/zackchou/Desktop/project _results")


```

```{r}
t <- newdata %>% 
  group_by(RequestType) %>% 
  summarise(count = n(),
         sumtime = sum(proc_time2),
         eff = count/sumtime) %>% 
  filter(!RequestType %in% c("Report Water Waste", "Other")) %>%
  arrange(-eff) %>% 
  ggplot(aes(x = reorder(RequestType, -eff), y = eff, fill = eff))+
  geom_bar(stat = "identity", show.legend = F)+
  coord_flip()+
  scale_fill_viridis(direction = 1)+
  labs(title = "Efficiency Rating", x = "Request Type", y = "Efficiency Ratio")+
  theme_classic()

ggsave("t.png",path = "/Users/zackchou/Desktop/project _results")

```

## low eff - homeless encampment
```{r}
u <- newdata %>% 
  filter(RequestType == "Homeless Encampment") %>% 
  group_by(RequestSource) %>% 
  summarise(pct = round(n()/14497, digit = 2)) %>% 
  arrange(-pct) %>% 
  slice(1:3) %>% 
  ggplot(aes(x = RequestSource, y = pct, fill = RequestSource))+
  geom_bar(stat = "identity", show.legend = F)+
  scale_fill_viridis(discrete = T)+
  labs(title = "Source of Homeless Encampment", x = "Request Sourece", y = "Percentage")+
  theme_classic()

ggsave("u.png",path = "/Users/zackchou/Desktop/project _results")


```

```{r}
v <- newdata %>% 
  filter(RequestType == "Homeless Encampment") %>% 
  group_by(RequestSource, wd, hr) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = wd, y = factor(hr), fill = count))+
  geom_tile()+
  scale_fill_viridis()+
  facet_wrap(~RequestSource)+
  labs(title = "Contact Method of Homeless Encampment",
       x = NULL, 
       y = NULL)+
  theme(axis.text.y = element_text(size = 6),
        axis.text.x = element_text(angle = 45))

ggsave("v.png",path = "/Users/zackchou/Desktop/project _results")
```

## low eff - multiple street light issue
```{r}
w <- newdata %>% 
  filter(RequestType == "Multiple Streetlight Issue") %>% 
  group_by(RequestSource) %>% 
  summarise(pct = round(n()/4989, digit = 2)) %>% 
  arrange(-pct) %>% 
  slice(1:3) %>% 
  ggplot(aes(x = RequestSource, y = pct, fill = RequestSource))+
  geom_bar(stat = "identity", show.legend = F)+
  scale_fill_viridis(discrete = T)+
  labs(title = "Source of Multiple Streetlight Issue", x = "Request Sourece", y = "Percentage")+
  theme_classic()
ggsave("w.png",path = "/Users/zackchou/Desktop/project _results")

```

```{r}
x <- newdata %>% 
  filter(RequestType == "Multiple Streetlight Issue") %>% 
  group_by(RequestSource, wd, hr) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = wd, y = factor(hr), fill = count))+
  geom_tile()+
  scale_fill_viridis()+
  facet_wrap(~RequestSource)+
  labs(title = "Contact Method of Multiple Streetlight Issue",
       x = NULL, 
       y = NULL)+
  theme(axis.text.y = element_text(size = 6),
        axis.text.x = element_text(angle = 45))
ggsave("x.png",path = "/Users/zackchou/Desktop/project _results")

```


## large request number - bulky item
```{r}
y <- newdata %>% 
  filter(RequestType == "Bulky Items") %>% 
  group_by(RequestSource) %>% 
  summarise(pct = round(n()/542372, digit = 2)) %>% 
  arrange(-pct) %>% 
  slice(1:3) %>% 
  ggplot(aes(x = RequestSource, y = pct, fill = RequestSource))+
  geom_bar(stat = "identity", show.legend = F)+
  scale_fill_viridis(discrete = T)+
  labs(title = "Source of Bulky Items", x = "Request Sourece", y = "Percentage")+
  theme_classic()
ggsave("y.png",path = "/Users/zackchou/Desktop/project _results")

```

```{r}
z <- newdata %>% 
  filter(RequestType == "Bulky Items") %>% 
  group_by(RequestSource, wd, hr) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = wd, y = factor(hr), fill = count))+
  geom_tile()+
  scale_fill_viridis()+
  facet_wrap(~RequestSource)+
  labs(title = "Contact Method of Bulky Items",
       x = NULL, 
       y = NULL)+
  theme(axis.text.y = element_text(size = 6),
        axis.text.x = element_text(angle = 45))

ggsave("z.png",path = "/Users/zackchou/Desktop/project _results")

```

## large request number - graffiti removal
```{r}
aa <- newdata %>% 
  filter(RequestType == "Graffiti Removal") %>% 
  group_by(RequestSource) %>% 
  summarise(pct = round(n()/279723, digit = 2)) %>% 
  arrange(-pct) %>% 
  slice(1:3) %>% 
  ggplot(aes(x = RequestSource, y = pct, fill = RequestSource))+
  geom_bar(stat = "identity", show.legend = F)+
  scale_fill_viridis(discrete = T)+
  labs(title = "Source of Graffiti Removal", x = "Request Sourece", y = "Percentage")+
  theme_classic()

ggsave("aa.png",path = "/Users/zackchou/Desktop/project _results")

```

```{r}
bb <- newdata %>% 
  filter(RequestType == "Graffiti Removal") %>% 
  group_by(RequestSource, wd, hr) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = wd, y = factor(hr), fill = count))+
  geom_tile()+
  scale_fill_viridis()+
  facet_wrap(~RequestSource)+
  labs(title = "Contact Method of Graffiti Removal",
       x = NULL, 
       y = NULL)+
  theme(axis.text.y = element_text(size = 6),
        axis.text.x = element_text(angle = 45))
ggsave("bb.png",path = "/Users/zackchou/Desktop/project _results")

```

## high eff - dead animal removal

```{r}
cc <- newdata %>% 
  filter(RequestType == "Dead Animal Removal") %>% 
  group_by(RequestSource) %>% 
  summarise(pct = round(n()/31697, digit = 2)) %>% 
  arrange(-pct) %>% 
  slice(1:3) %>% 
  ggplot(aes(x = RequestSource, y = pct, fill = RequestSource))+
  geom_bar(stat = "identity", show.legend = F)+
  scale_fill_viridis(discrete = T)+
  labs(title = "Source of Dead Animal Removal", x = "Request Sourece", y = "Percentage")+
  theme_classic()

ggsave("cc.png",path = "/Users/zackchou/Desktop/project _results")

```

```{r}
dd <- newdata %>% 
  filter(RequestType == "Dead Animal Removal") %>% 
  group_by(RequestSource, wd, hr) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = wd, y = factor(hr), fill = count))+
  geom_tile()+
  scale_fill_viridis()+
  facet_wrap(~RequestSource)+
  labs(title = "Contact Method of Dead Animal Removal",
       x = NULL, 
       y = NULL)+
  theme(axis.text.y = element_text(size = 6),
        axis.text.x = element_text(angle = 45))

ggsave("dd.png",path = "/Users/zackchou/Desktop/project _results")

```



```{r}
s <- newdata %>% 
  group_by(RequestType, wd, hr, RequestSource) %>% 
  summarise(avg = mean(proc_time)) %>% 
  ggplot(aes(x = factor(hr), y = avg, fill = RequestSource))+
  geom_bar(stat = "identity", position = "dodge", show.legend = F)+
  facet_wrap(~RequestSource)+
  theme_minimal()+
  labs(title = "Avg. Procedure Time via Each Request Source in 24 Hours", x = "Hour in a Day", y = "Procedure Days")+
  theme(axis.text.y = element_text(size = 4))+
  scale_fill_viridis(discrete = T)+
  coord_flip()

ggsave("s.png",path = "/Users/zackchou/Desktop/project _results")

```

```{r}
ll <- newdata %>% 
  filter(RequestSource %in% c("Mobile App")) %>% 
  group_by(RequestType) %>% 
  summarise(count = n()) %>% 
  arrange(-count) %>% 
  ggplot(aes(x = reorder(RequestType, count), y = count, fill = count))+
  geom_bar(stat = "identity", show.legend = F)+
  scale_fill_viridis()+
  coord_flip()+
  theme_classic()+
  labs(title = "Ranking of Request Type via Mobile App", x = "Request Type", y = "Count")

ggsave("ll.png",path = "/Users/zackchou/Desktop/project _results")
```

## 1. overall time dist data1
## 2. overall spatial

```{r}
dd <- call1 %>%
        filter(!is.na(hr)) %>% 
        group_by(wd, hr) %>% 
        summarise(count = n()) %>% 
        ggplot(aes(x = wd, y = factor(hr), fill = count))+
        geom_tile()+
        scale_fill_viridis()+
        labs(title = "Overall Call Time Distribution 2011-2015", y = "Hour", x = "Weekday", fill = "Frequency")+
        theme_classic()+
        theme(panel.background = element_rect(color = "black"), 
              axis.text.x = element_text(angle = 45, vjust = 0.5),
              axis.text.y = element_text(size = 6))
        
ggsave("dd.png", path = "/Users/zackchou/Desktop/project _results")
```

```{r}
ff <- la+
  geom_polygon(data = joindata1, 
               aes(x = long, y = lat, group = group, fill = count, alpha = count),
               color = "darkgrey")+
  theme_void()+
  scale_fill_viridis()+
  labs(title = "Overall Call Frequency by Zip Area 2011-2015", fill =  "Frequency", alpha = "Transparency")

ggsave("ff.png", path = "/Users/zackchou/Desktop/project _results")

```

```{r}


gg <- newdata %>% 
  filter(ZipCode == 90011) %>% 
  group_by(wd, hr) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = wd, y = factor(hr), fill = count))+
  geom_tile()+
  scale_fill_viridis()+
  theme_classic()+
  labs(title = "Overall Time Distribution of 'Zip Code 90011'",
       y = "Hour", 
       x = "Weekday")+
  theme(axis.text.y = element_text(size = 6),
        axis.text.x = element_text(angle = 45))

ggsave("gg.png", path = "/Users/zackchou/Desktop/project _results")


```

```{r}
f1 <- newdata %>% 
  group_by(RequestSource) %>% 
  summarise(count = n(), 
            pct = round(count/1099501, digit = 2)) %>% 
  arrange(-count)
sum(f1$count)
```


```{r}
range(newdata$CreatedDate)

newdata <- newdata %>% 
  mutate(date = str_sub(CreatedDate, end = 10))


ss <- call1 %>% 
  group_by(Call.Type) %>% 
  summarise(count = n()) %>% 
  arrange(-count) %>% 
  ggplot(aes(x = reorder(Call.Type, -count), y = count, fill = count))+
  geom_bar(stat = "identity", show.legend = F)+
  theme_classic()+
  labs(title = "Call Resolution Type 2011-2015", x = "Call Resolution", y = "Count")+
  scale_fill_viridis()
  
ggsave("ss.png", path = "/Users/zackchou/Desktop/project _results")
```






